---
title: "fdic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fdic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r setup}
library(fdic)
```

Each function in {fdic} accepts the following arguments:

- `api_key`: Your FDIC API key
- `filters`: One or more filters to apply when requesting data using [Elasticsearch Query String Syntax](https://www.elastic.co/docs/reference/query-languages/query-dsl/query-dsl-query-string-query#query-string-syntax)
- `fields`: One or more fields to include in the response
- `sort_by`: A field name to sort the response by
- `descending`: A flag to specify the direction to `sort_by` (if `sort_by` is specified)
- `limit`: The number of records to include in the response (up to a maximum of 10,000)

While most of the arguments are relatively straightforward, there are some idiosyncrasies with both the `fields` and `filters` arguments that are worth discussing.

### Available API Fields

{fdic} contains eight internal datasets documenting the [current API endpoint definition files](https://api.fdic.gov/banks/docs/) provided by the FDIC. Each dataset corresponds to one of the functions contained in {fdic} and is named by prefixing the endpoint with `fdic_` (e.g., `fdic_institutions` for `get_institutions()`). Each dataset can be accessed directly by name, as demonstrated below.

```{r example-1}
# Dropping `description` for example due to length of field
head(fdic_locations) |>
  subset(select = -description)
```

During package development, it was noted that *most* fields returned by the API are documented in these internal datasets. **However, there are several instances where fields are either no longer available or new (undocumented) fields have been added.**

{fdic} functions evaluate the values supplied to the `fields` argument and will raise a warning if a field is not returned in the response. However, it can be helpful to call an {fdic} function with no `fields` argument and `limit = 1` to return the current endpoint definition, as demonstrated below:

```{r example-2}
# Review current endpoint definition
get_locations(limit = 1) |>
  names()
```

Alternatively, the BankFind Suite offers a [Glossary and Variable Definition](https://banks.data.fdic.gov/bankfind-suite/help?helpTopic=glossary-and-variable-definitions) table which may provide more up-to-date information on API fields.

By familiarizing yourself with the available `fields`, you can begin to refine your {fdic} queries by passing `filters` to target the data you are most concerned with. The next section provides a brief primer on the `filters` syntax.

### Filtering Using Elasticsearch Query String Syntax

The [FDIC Bank Suite API](https://api.fdic.gov/banks/docs/) uses [Elasticsearch Query String Syntax](https://www.elastic.co/docs/reference/query-languages/query-dsl/query-dsl-query-string-query#query-string-syntax) to filter results.

Elasticsearch Query String Syntax is a mini-language that allows for a customized search of the data, using familiar terms and operators to facilitate the filtering.

By passing a valid Elasticsearch Query String to the `filters` argument of an {fdic} function, you can conveniently manipulate the data provided in response.

The following examples demonstrate several ways to use Elasticsearch Query Strings in {fdic} functions to collect the data of interest.

```{r example-3}
# Search for five active institutions in New York
# Return all available fields
get_institutions(
  filters = "STALP:NY AND ACTIVE:1",
  limit = 5
)
```

```{r example-4}
# Collect location data for five branches of a specific institution
# Return all available fields
get_locations(
  filters = "CERT:33124",
  limit = 5
)
```

```{r example-5}
# Explore the 2025 Summary of Deposit data for non-community banks in New York
# Collect the coordinates for the top five branch locations by total deposits
get_sod(
  filters = "STALP:NY AND !(CB:1) AND YEAR:2025",
  fields = c("DEPSUM", "NAMEBR", "SIMS_LATITUDE", "SIMS_LONGITUDE", "YEAR"),
  sort_by = "DEPSUM",
  descending = TRUE,
  limit = 5
)
```
